---
title: "FCC Broadband Deployment Data Processing"
author: "Cary K. Jim"
date: 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "D://Documents//R//Dissertation")
```

Data Source from FCC Fixed Broadband Deployment 
<https://broadbandmap.fcc.gov/#/data-download>

Background Information:
Broadband Technologies can be categorized as:
- ADSL
- Cable
- Fiber
- Fixed wireless 
- Satellite
- Other

Speed category of download/Upload (Mbps: Megabits per second) in more than > or equal = to:
- 0.2/0.2
- 4/1
- 10/1
- 25/3
- 100/10
- 250/25
- 1000/100


```{r cars}
library(tidyverse)
library(dplyr)
library(readr)
library(stringr)
library(janitor)
library(readxl)
```
The Fixed Broadband Deployment Data from FCC Form 477 included All, FIXED, and MOBILE broadband providers.

The Explanation of the Data <https://www.fcc.gov/general/explanation-broadband-deployment-data> provides the description of each element. 

Definition of ISP <https://www.law.cornell.edu/wex/internet_service_provider_(isp)>

Instruction for ISP on Reporting on 477 Form <https://us-fcc.app.box.com/v/FBDFormatting>

# FCC data without satellite from June 2020 (~ 4.76 GB)

```{r}
fcc_none_satellite <- read.csv("./Datasets/FCC/fbd_us_without_satellite_jun2020_v1.csv", header = T)
```

Varibles that can be useful:
BlockCode - 15 digit census block code used in the 2010 US Census (referred as GEOID)
<https://transition.fcc.gov/form477/Geo/more_about_census_blocks.pdf>
TechCode - is a 2 digita category code for the type of broadband technology <https://www.fcc.gov/general/technology-codes-used-fixed-broadband-deployment-data>
Consumer - binary indicator of the ISP offer service for residential consumer
MaxAdDown - Max. advertised downstream speed/bandwidth by provider for the block customer
MasAdUp = Max. advertised upstream speed/bandwidth by provider for the block customer
Business - binary indicator of the ISP offer service for business/government consumer (as of Dec 2019, business speed data are no longer collected)

Notes on the dataset per FCC documentation: 
- Records submitted for any water-only census blocks were removed.

## Data Characteristics 
```{r}
colnames(fcc_none_satellite)
```

```{r}
str(fcc_none_satellite)
```

```{r}
head(fcc_none_satellite)
```

```{r}
#Convert the block code as character for parsing 
fcc_none_satellite[10] <- as.character(fcc_none_satellite$BlockCode)
```
View the dataframe again 
```{r}
head(fcc_none_satellite)
```
Census Block by State in FCC site <https://www.fcc.gov/general/census-blocks-state>

Check how many census block group is there.
```{r}
fcc_none_satellite %>% group_by(BlockCode) %>% summarize()
```

substring in base R
```{r}
splitted <- t(sapply(fcc_none_satellite$BlockCode, function(x) substring(x, first=c(1,2,4), last=c(1,3,6))))
```

cbind(df, splitted)
```{r}
#Getting a list of BlockCode 
block <- as.data.frame (fcc_none_satellite[, 10])
block %>% n_distinct() #Check for distinct block in dataframe, there are 9527112
```
```{r}
colnames(block)
```

```{r}
splitted <- substr(block$"fcc_none_satellite[, 10]", 1, 11)
head(splitted)
```

```{r}
#Select the columns that is useful 
```


# FCC data with satellite from June 2020 (~ 10.4GB)

```{r}
fcc_w_satellite <- read.csv("D:\\Documents\\R\\Dissertation\\Datasets\\FCC\\fbd_us_with_satellite_jun2020_v1.csv",header = T)
```




