---
title: "Microsoft Broadband Usage 2020 Data Processing"
author: "Cary K. Jim"
output:
  github_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 80
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r  Library}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(janitor)
library(psych)
library(readxl)
library(psychTools)
```

# Microsoft Broadband Usage Data
Data Source:<https://github.com/microsoft/USBroadbandUsagePercentages>

# Oct 2020 data

## County Level

```{r}
microsoft_2020_county <- 'https://raw.githubusercontent.com/microsoft/USBroadbandUsagePercentages/master/dataset/broadband_data_2020October.csv'

# The regular reading will not work because of in-line text within the csv file in the 2020 dataset
usage_county_2020 <- read.csv(url(microsoft_2020_county), header = T, sep = ",", strip.white = TRUE, fileEncoding="UTF-8-BOM", na.strings = "-")

#' reload the usage data by skipping the top 18 rows (n = 3142)
usage_county_2020 <- read.csv(url(microsoft_2020_county), skip = 18, header = T, sep = ",", strip.white = TRUE, fileEncoding="UTF-8-BOM")

```

### Dataframe Overview

```{r}
head(usage_county_2020)
```

```{r}
str(usage_county_2020)
```

Notice a few things from looking at the information. COUNTY.ID was
loaded as integer,'BROADBAND.AVAILABILITY.PER.FCC' & 'BROADBAND.USAGE'
was loaded as characters. We have to correct the data types before next
step.

```{r Rename Columns}
#Rename columns to make time frame explicit in the variables 
usage_county_2020 <- usage_county_2020 %>%
  rename(FCC2019_AVAIL = BROADBAND.AVAILABILITY.PER.FCC,
         USAGE_2020 = BROADBAND.USAGE)
```

```{r Convert to Numeric}
#Convert the columns to number if read in as characters
usage_county_2020[4:5] <- lapply(usage_county_2020[4:5], as.numeric)
```

```{r Convert to Characters}
#Then, convert the COUNTY.ID to character 
usage_county_2020[2] <- as.character(usage_county_2020$COUNTY.ID)
# For a proper FIPS, there should be leading zero to it. For now, I'm ignoring this matter and will handle it later.
```

### Review data type after change

```{r}
str(usage_county_2020)
```

### Missing values in the Oct 2020 County Level Usage Dataset

```{r}
#There are 20 County without either Broadband Availability information or Usage Information 
usage_county_2020[!complete.cases(usage_county_2020), ]
```

### Number of Countes by State

```{r}
usage_county_2020 %>% count(ST)
```

## ZIP Code Level

```{r}
microsoft_2020_ZIP <- "https://raw.githubusercontent.com/microsoft/USBroadbandUsagePercentages/master/dataset/broadband_data_zipcode.csv"
usage_ZIP_2020 <- read.csv(url(microsoft_2020_ZIP), header = T, sep = ",", strip.white = TRUE, fileEncoding="UTF-8-BOM")
```

### Dataframe Overview

```{r}
head(usage_ZIP_2020)
```

```{r Rename_Columns}
#Rename columns to make time frame explicit in the variables and fixed the long names 
usage_ZIP_2020 <- usage_ZIP_2020 %>% 
  rename(USAGE_2020 = BROADBAND.USAGE)

colnames(usage_ZIP_2020)[6:7] <- c("MAE", "Percentile")
```

### Review data types

```{r}
str(usage_ZIP_2020)
```

For the ZIP level dataset, the usage information and its error range
were in the correct form as numeric. For COUNTY.ID and POSTAL.CODE, they
need to be converted to characters. Also, POSTAL.CODE will be renamed to
ZIP for ease of use at a later time.

```{r}
usage_ZIP_2020[3:4] <- lapply(usage_ZIP_2020[3:4], as.character)
```

Rename the POSTAL.CODE column to ZIP

```{r Rename Postal Code to ZIP}
colnames(usage_ZIP_2020)[4] <- "ZIP"
```

### Review data type after conversion

```{r}
str(usage_ZIP_2020)
```

### Check Missing Values

```{r Missing Values}
usage_ZIP_2020[!complete.cases(usage_ZIP_2020), ]
```

### Check Duplicated Rows

```{r Duplicated ZIP code}
sum(duplicated(usage_ZIP_2020$ZIP))
```

```{r Which ZIP code is duplicated?}
usage_ZIP_2020$ZIP[duplicated(usage_ZIP_2020$ZIP)]
```

```{r Example Kodiak Island Borough, AK}
usage_ZIP_2020 %>% filter(ZIP == "99554" | ZIP == "99658") 
# This county has two ID in which the 2150 is the correct COUNTY.ID as of 2010. County.ID 2158 is an old COUNTY.ID. 
```
Reference to Census Technical Documentation on County Changes in 2010.
<https://www.census.gov/programs-surveys/geography/technical-documentation/county-changes.2010.html>
```{r Remove the old COUNTY.ID for Kodiak}
usage_ZIP_2020 <- usage_ZIP_2020[!(usage_ZIP_2020$COUNTY.ID == "2158"),]
# The resulting dataframe is 32718 rows 
```
### Re-check duplicated rows after handling the Kodiak Island Borough records 
```{r}
usage_ZIP_2020$ZIP[duplicated(usage_ZIP_2020$ZIP)]
```
```{r Moody County, SD has duplicates}
usage_ZIP_2020 %>% filter(ZIP == "57065" | ZIP == "57017") 
```
Moody County is not listed in the County Name Change webpage on Census. The U.S. Census Quick Facts is used to check the current County ID for Moody, in SD. <https://www.census.gov/quickfacts/fact/table/moodycountysouthdakota,US/PST045221>
The correct County ID for Moody, SD is 46101. 
```{r Remove the old COUNTY.ID for Moody}
usage_ZIP_2020 <- usage_ZIP_2020[!(usage_ZIP_2020$COUNTY.ID == "46102"),]
# The resulting dataframe is 32707 rows 
```
### Re-check duplicated rows after handling the Moody County records 
```{r}
usage_ZIP_2020$ZIP[duplicated(usage_ZIP_2020$ZIP)]
```
There is still 54 rows of records that has duplicated county by ZIP. Therefore, a subset of the dataframe is used. 
```{r}
subset(usage_ZIP_2020,duplicated(ZIP))
```
There are two more counties that have similar issues.   
```{r Todd County, MN}
usage_ZIP_2020 %>% filter(ZIP == "56534" | ZIP == "56437") 
```
Todd County, MN should be COUNTY.ID == 27153
```{r Washington County, MN}
usage_ZIP_2020 %>% filter(ZIP == "55082" | ZIP == "56160") 
```
Washington County, MN should be 27163
```{r Remove Todd and Washington County in MN}
usage_ZIP_2020 <- usage_ZIP_2020[!(usage_ZIP_2020$COUNTY.ID == "27111"),]
# The resulting dataframe is 32677 rows after removing Todd duplicates
usage_ZIP_2020 <- usage_ZIP_2020[!(usage_ZIP_2020$COUNTY.ID == "27165"),]
# The resulting dataframe is 32653 rows after removing Washington County duplicates
```
### Re-check duplicated record 
```{r}
usage_ZIP_2020$ZIP[duplicated(usage_ZIP_2020$ZIP)]
```
The duplicated records have been removed. The distinct rows of ZIP code is 32653 in the ZIP 2020 usage data file. 

### Counts of County and State by ZIP
```{r ZIP code by each State}
usage_ZIP_2020 %>% count(ST)
```

```{r}
usage_ZIP_2020 %>% count(COUNTY.ID)
```

```{r ZIP by County by State}
#Each row is a ZIP code and so counting the COUNTY.NAME will show how many ZIP is in each county in each State.
usage_ZIP_2020 %>% group_by(ST) %>% count(COUNTY.NAME)
```

```{r Unique count of ZIP code}
length(unique(usage_ZIP_2020$ZIP))
```
```{r Export file, include=FALSE}
write.csv(usage_county_2020, "./Processed_files/usage_county_2020.csv", row.names = F)
write.csv(usage_ZIP_2020, "./Processed_files/usage_ZIP_2020.csv", row.names = F)
```


